---
title: "Recursive Multivariate Testing for Cohort Clustering with Disparate Covariate Information"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Recursive Multivariate Testing for Cohort Clustering with Disparate Covariate Information}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Load packages
```{r setup}
suppressMessages({
  library(dendextend)
  library(randomForestSRC)
  library(SHT)
  library(mice)
  library(stringr)
  library(relate)
})

```


## Data setup
```{r}
set.seed(123)
#load simulated dataset from the package
data("cohort_na_df")
cohort_na_df[199:202,] 
#save the true distribution
true_distribution <- cohort_na_df$distribution 
cohort_na_df <- cohort_na_df %>% dplyr::select(-distribution)
# examine missing pattern (e.g., some cohorts did not collect specific information)
cohort_na_df %>% find_na_level(clusterid = "cohortid")
#add random missingness in addition to the systematic missingness for a more realistic scenario
random_missing_row <- sample(1:nrow(cohort_na_df), floor(nrow(cohort_na_df)/2))
missing_col <- c(1,8,13,14)
cohort_na_df[random_missing_row, missing_col] <- NA
# examine missing pattern after adding random missingness
cohort_na_df %>% find_na_level(clusterid = "cohortid")
```

## Conduct unsupervised random forest using cohort id as the outcome
```{r rf}
analytic.rfsrc <- rfsrc.fast(cohortid ~ ., #regress cohort id on all other variables 
                             data = cohort_na_df,
                             distance=T, ntree=1500, nsplit=3,
                             na.action="na.impute",nimpute=1) # impute within nodes
distance.matrix <- analytic.rfsrc$distance     # extract distance matrix 
```

## Hierachical clustering of the average distance matrix
```{r h-cluster}
#calculate average distance matrix
dist_avg_mat <- avg_dist_func(distance.matrix, cohort_na_df$cohortid)

#cluster average distance matrix and 
#identify which cohorts are clustering together on dendogram
dmean <- as.dist(dist_avg_mat)

#hierarchical clustering 
hc2 <- hclust(dmean, method = "ward.D2")
dend <- as.dendrogram(hc2)
plot(dend)
table(true_distribution, cohort_na_df$cohortid, dnn = c("True Cluster","Cohort ID"))
```

## Within cohort imputation to account for Missing At Random (MAR) variables 
We generally do not impute the missing variables that are completely missing or 
missing more than 50%.

The `relate::mice_group_impute` in this package provides imputation by group regardless of missingness level (imputeLevel = 1). 
Please find `miceadds::mice.impute.bygroup` for similar functionality. 
```{r groupimp, eval=F}
mice_group_impute(df = cohort_na_df, clusterid = "cohortid", imputeLevel = 1, miceArg = list(method = 'mean', maxit = 1))
```

## Recursive BG testing based on the previous dendrogram 
```{r recursion, warning=F, message=F, error=F}
# we use mean imputation and asymptotic BG test for illustrative purposes. 
ll <- recursive.test(dend = dend, df = cohort_na_df, cohortid.var = "cohortid", 
           verbose = T, saveIntermediate = F, 
           BG.method = 'permutation', n_perm = 100, impute = T, 
           miceArgs = list(method = c("mean"), m =1, maxit = 5, seed = 123, printFlag = F))

ll$clusters
table(true_distribution, cohort_na_df$cohortid, dnn = c("True Cluster","Cohort ID"))
```



